---
title: "Raw Data Report"
execute:
  echo: false
  warning: false
  error: false
format: html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
options(tidyverse.quiet = T)
```


```{r}
#| echo: false
library(targets)
library(ggplot2)
```

## (Not really all that) Quick & Dirty

-   These are the raw data results for *most* of the variables for chapter 1 data
-   NOTES: still missing tree diversity measures
-   still missing street scale data (something weird happening there)

## Points of Discussion

-   Comparable street segments
-   Tree size skewness measurements
-   Tree species richness + iNEXT
-   CMA vs weighted census tracts
-   Montreal DBHs aka what is happening here
-   DSM correction (I think I need to correct for sea level)

## Workflow

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 15
#| fig-height: 15
#| output: false

tar_visnetwork(outdated = F)
```

## Tree Stuff

Number of trees per city:

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 15
#| fig-height: 15
tar_read(tree_count_city)
```

Number of tree **species** per city:

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 15
#| fig-height: 15
tar_read(sp_count_city)
```

Tree size distribution per city:

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 15
#| fig-height: 15
tar_read(skewness_vis_city)
```

## City Scale (CMA level)

### Response Variables

-   Response variables are proxies for ES
    -   Land Surface Temperature
    -   Air Pollution measures for 6 pollutants

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 15
#| fig-height: 15
tar_read(city_response_facet)
```

### Independent Variables - Trees

-   currently missing diversity measures but have included tree size and tree density

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 15
#| fig-height: 15
tar_read(city_tree_facet)
```

### Independent Variables - Built Infrastructure

-   includes building density, building height, road density, proportion of road types, NDVI, and NDBI

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 15
#| fig-height: 15
tar_read(city_build_facet)
```

### Independent Variables - Census

-   includes data from 2021 census on:
    -   totpop = Population 2021

    -   popdens = Pop density per sq km

    -   area = Land area sq km

    -   sidho = Single-detached house

    -   semho = Semi-detached house

    -   rowhou = Row house

    -   aptdup = Apartment or flat in a duplex

    -   aptbui = Apartment in a building that has fewer than five storeys

    -   aptfiv = Apartment in a building that has five or more storeys otsiho = Other single-attached house

    -   mvdwel = Movable dwelling (9)

    -   medinc = Median after-tax income in 2015 among recipients

    -   lowinc = Prevalence of low income based on the Low-income measure, after tax (LIM-AT) (%)

    -   recimm = Total - Immigrant status and period of immigration for the population in private households - 25% sample data (63) -\> Immigrants -\> 2011-2016 (recent immigrants)

    -   aborig = Total - Aboriginal identity for the population in private households - 25% sample data (104) -\> Aboriginal identity (105)

    -   vismin = Total - Visible minority for the population in private households - 25% sample data (121) -\> Total visible minority population (122)

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 15
#| fig-height: 15
tar_read(city_census_facet)
```


## Neighbourhood Scale

### Response Variables

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 15
#| fig-height: 15
tar_read(neighbourhood_response_facet)
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 15
#| fig-height: 15
tar_read(neighbourhood_response_pca)
```

### Independent Variables - Trees

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 15
#| fig-height: 15
tar_read(neighbourhood_tree_facet)
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 15
#| fig-height: 15
tar_read(neighbourhood_tree_pca)
```

### Independent Variables - Built Infrastructure

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 15
#| fig-height: 15
tar_read(neighbourhood_build_facet)
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 15
#| fig-height: 15
tar_read(neighbourhood_build_pca)
```

### Independent Variables - Census

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 15
#| fig-height: 15
tar_read(neighbourhood_census_facet)
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 15
#| fig-height: 15
tar_read(neighbourhood_census_pca)
```

## Street Scale

To come...